Stochastic Variance-Reduced Policy Gradient

### 2018-02-07 svrpg.tex

We can now introduce Stochastic Variance-Reduced Policy Gradient (SVRPG) for a generic policy gradient estimator $g$ (see Eq.~\eqref{E:policygradient.estimate}).
% For simplicity, we focus on the REINFORCE estimator but the extension to G(PO)MDP is straightforward.
Between snapshots, the SVRPG estimate is corrected using importance weighting, for any $t \in \{0,\ldots,m-1\}$
\begin{align*}
        \blacktriangledown J(\vtheta_{t}) &= \wh{\nabla}_N J(\wt{\vtheta}) + g(\tau_i|\vtheta_t) - \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta}),
\end{align*}
% \begin{align*}
% 	\blacktriangledown J(\vtheta) &= \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \vert \tilde{\vtheta})\Reward(\tau_j) \\
% 		&+ \nabla\log\pi(\tau_i \vert \vtheta)\Reward(\tau_i) 
% 		- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \tilde{\vtheta})\Reward(\tau_i),
% \end{align*}
where $\wh{\nabla}_N J(\wt{\vtheta})$ is as in Eq.~\eqref{E:policygradient.estimate} where $\mathcal{D}_N$ is sampled using the snapshot policy $\pi_{\wt{\vtheta}}$, $\tau_i$ is sampled from the current policy $\pi_{\vtheta_t}$, and $\omega(\tau|\vtheta_t, \wt{\vtheta}) = \frac{p(\tau|\wt{\vtheta})}{p(\tau|\vtheta_t)}$ is an importance weight from $\pi_{\vtheta_t}$ to the snapshot policy $\pi_{\wt{\vtheta}}$. 
Similarly to SVRG, we have that $\vtheta_0 := \wt{\vtheta}$, and the update is a FG step.
Our update is still fundamentally on-policy since the weighting concerns only the correction term. However, this partial ``off-policyness'' is enough to introduce variance. This is a well-known issue of importance sampling~\citep[\eg][]{thomas2015high}. To mitigate it, we use mini-batches of trajectories of size $B \ll N$ to average the correction, \ie
\begin{align}\label{E:svrpg.estimate.batch}
        \blacktriangledown J(\vtheta_{t}) &:= v_t= \wh{\nabla}_N J(\wt{\vtheta})\\ \notag
                                            & \quad{} + 
        \underbracket{
        \frac{1}{B} \sum_{i=0}^{B-1} \left[
        g(\tau_i|\vtheta_t) - \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
\right]}_{c_t}.
\end{align}
We want just to stress that $\EVV[\tau_i \sim p(\cdot|\vtheta_t)]{\frac{1}{B} \sum_{i=0}^{B-1} \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})} = \nabla J(\wt{\vtheta})$.\footnote{The reader can refer to Appendix~\ref{} for off-policy gradients and variants of REINFORCE and G(PO)MDP.}
This property will be used to prove Lemma~\ref{L:svrpg.properties}.
%\begin{align*}
%&\blacktriangledown J(\vtheta) = \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \vert \tilde{\vtheta})\Reward(\tau_j) +\\
%&\quad\frac{1}{B}\left[\sum_{i=0}^{B-1}
%\nabla\log\pi(\tau_i \vert \vtheta)\Reward(\tau_i) 
%- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \tilde{\vtheta})\Reward(\tau_i)\right].
%\end{align*}
The use of mini-batches is a common practice in SVRG since it can yield a performance improvement even in the supervised case~\citep{harikandeh2015stopwasting,konevcny2016mini}. It is easy to show that the SVRPG estimator has the following, desirable properties (see Appendix~\ref{app:proofs}):
\begin{lemma}\label{L:svrpg.properties} \todopirout{extended to any unbiased estimator}
        {\color{blue} Let $\wh{\nabla}_N J(\vtheta)$ be an unbiased estimator of~\eqref{E:policygradient} }
%Denote by $\wh{\nabla}_N^{\textsc{RF}} J(\vtheta)$ the REINFORCE estimator in~\eqref{E:policygradient.estimate} 
and let $\vtheta^* \in \argmin_{\vtheta} \{J(\vtheta)\}$. Then, the SVRG estimate in~\eqref{E:svrpg.estimate.batch} is \emph{unbiased}
\[
\mathop{\mathbb{E}}
\left[\blacktriangledown J(\vtheta)\right] = \gradJ{\vtheta}.
\]
and regardless of the mini-batch size $B$:\footnote{
For any vector $\mathbf{x}$, we use $\Var[\mathbf{x}]$ to denote the trace of the covariance matrix, \ie $\Tr\EVV[]{(\mathbf{x}-\EVV[]{\mathbf{x}})(\mathbf{x}-\EVV[]{\mathbf{x}})^T}$.}
\[
	\Var\left[\gradBlack{\vtheta^*}\right] = 
    \Var\left[\wh{\nabla}_N J(\vtheta^*)\right].
\]
\end{lemma}

@@

\begin{algorithm}[tb]
        \caption{SVRPG}
	\label{alg:svrpg}
	\begin{algorithmic}
        \STATE {\bfseries Input:} number of epochs $S$, epoch size $m$, step size $\alpha$, batch size $N$, mini-batch size $B$, gradient estimator $g$, initial parameter $\vtheta_{0}^m := \wt{\vtheta}^0$
		\FOR{$s=0$ {\bfseries to} $S-1$}
        \STATE $\vtheta_0^{s+1} := \tilde{\vtheta}^{s} = \vtheta_{m}^s$
        \STATE Sample $N$ trajectories $\{\tau_j\}$ from $p(\cdot\vert\tilde{\vtheta}^{s})$
        \STATE $ \wt{\mu} = \gradApp{\wt{\vtheta}^{s}}{N}$ (see Eq.~\eqref{E:policygradient.estimate})% = \frac{1}{N}\sum_{j=0}^{N-1}g(\tau_j | \wt{\vtheta}^s)$
		\FOR{$t=0$ {\bfseries to} $m-1$}
        \STATE Sample $B$ trajectories $\{\tau_i\}$ from $p(\cdot\vert\vtheta_t^{s+1})$
%        \STATE $c^{s+1}_t = \frac{1}{B} \sum\limits_{i=0}^{B-1} \left( g(\tau_i|\vtheta_t^{s+1}) - \omega(\tau_i|\vtheta^{s+1}_t, \wt{\vtheta}^s) g(\tau_i| \wt{\vtheta}^s) \right)$
        \STATE $c^{s+1}_t = \frac{1}{B} \sum\limits_{i=0}^{B-1}
        \begin{aligned}[t]
                \Big( & g(\tau_i|\vtheta_t^{s+1})\\ 
                      & \quad{} - \omega(\tau_i|\vtheta^{s+1}_t, \wt{\vtheta}^s) g(\tau_i| \wt{\vtheta}^s) \Big)
        \end{aligned}$
%        \STATE $\begin{aligned}[t]
%                &v^{s+1}_t = \wt{\mu}\\ 
%                & +\frac{1}{B} \sum\limits_{i=0}^{B-1} \left( g(\tau_i|\vtheta_t^{s+1}) - \omega(\tau_i|\vtheta^{s+1}_t, \wt{\vtheta}^s) g(\tau_i| \wt{\vtheta}^s) \right)
%                \end{aligned}$
        \STATE $v^{s+1}_t = \wt{\mu} + c^{s+1}_t$ % \gradApp{\wt{\vtheta}^{s}}{N} + c^{s}_t$
%%		\begin{align*}
%%		\blacktriangledown J(\vtheta_t^{s+1}) = 
%%		&\gradApp{\tilde{\vtheta}^s}{N} \\
%%		&+\frac{1}{B}\sum_{i=0}^{B-1}\left[ 
%%		\score{}{\tau_i\vert\vtheta_t^{s+1}}\Reward(\tau_i)\right. \\
%%		&\left. - \omega(\tau_i)\score{}{\tau_i \vert \tilde{\vtheta}^{s}}\Reward(\tau_i)\right]
%%		\end{align*}
        \STATE $\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha v^{s+1}_t$
		\ENDFOR
		\ENDFOR
        \STATE {\bfseries return} $\vtheta_t^{s+1}$ with $(s,t)$ picked uniformly at random from $\{[0,S-1]\times[0,m-1]\}$
	\end{algorithmic}
\end{algorithm}

@@

\subsection{Baseline}
As reported in Section~\ref{subsec:PolicyGradient}, a popular variance reduction tool in Reinforcement Learning is the employment of baselines in policy gradient estimation.
This tool is orthogonal to the variance reduction technique proposed in this paper. In Section~\ref{sec:exp} we try to combine SVRPG with a non-stationary, state-dependent, learned linear baseline like the one used in~\cite{duan2016benchmarking}, with the variant that baseline parameters are updated only at the snapshot, with trajectories sampled from $\pi_{\tilde{\vtheta}^s}$.

@@

+\frac{L_2^2}{B}\Ets[t-1]{\norm[2]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}

@@

	= \Ets{\norm[]{\gradApp{\tilde{\vtheta}^s}{N}

@@

    \omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)}^2} \nonumber\\

@@

	&= \mathbb{E}_{t\vert s}\left[\left\|\gradApp{\tilde{\vtheta}^s}{N}

@@

			\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)\right)

@@

            \pm \gradJ{\vtheta_t^{s+1}} \pm \gradJ{\tilde{\vtheta}^s} 
			% -\gradJ{\vtheta_t^{s+1}} + \gradJ{\tilde{\vtheta}^s}
			% +\gradJ{\vtheta_t^{s+1}} - \gradJ{\tilde{\vtheta}^s}

@@

	+\Es{\norm[]{\gradApp{\tilde{\vtheta}^s}{N} - \Es{\gradApp{\tilde{\vtheta}^s}{N}}}^2} \nonumber\\

@@

				\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)\right)}\right\|^2\right] 

@@

@@

			\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)\right)}^2} \label{eq:2}\\

@@

    \omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s) \pm g(\tau_i|\wt{\vtheta}^s) }^2} \nonumber\\

@@

%%	+\frac{1}{N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)}

@@

%%			-g(\tau_i\vert\tilde{\vtheta}^s)
%%			+g(\tau_i\vert\tilde{\vtheta}^s) 
%%			-\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)}^2} \nonumber\\

@@

			-g(\tau_i\vert\tilde{\vtheta}^s)}^2} 

@@

			\Ets{\norm[]{g(\tau_i\vert\tilde{\vtheta}^s) 
			-\omega(\tau_i)g(\tau_i\vert\tilde{\vtheta}^s)}^2} \nonumber\\

@@

    +\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}

@@

			-\omega(\tau_i))g(\tau_i\vert\tilde{\vtheta}^s)}^2} \label{eq:3}\\

@@

	&\qquad+\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\tilde{\vtheta}^s}^2}

@@

%Lemma 0
\begin{restatable}[]{lemma}{auxzero}\label{lemma:aux0}
Under Assumption \ref{ass:bounded_score}, for any function $\varphi(\vtheta_t^{s+1})$ which is deterministic for a fixed $\vtheta_t^{s+1}$:
\begin{align*}
\left|\Ets[t]{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}}
-\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}}
\right|
\leq
\frac{1}{2N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{\varphi(\vtheta_t^{s+1})}^2}
\end{align*}
\end{restatable}
\begin{proof}
\begin{align}
	\Ets{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}}
	&=
	\Ets{\dotprod{\gradIdeal{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}} +
	\Ets[t-1]{\dotprod{e_s}{\varphi(\vtheta_t^{s+1})}} \label{eq:6}\\
	&=
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}} +
	\Ets[t-1]{\dotprod{e_s}{\varphi(\vtheta_t^{s+1})}} \label{eq:7}\\
	&=
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}} \nonumber\\
	&\qquad+
	\dotprod{\Ets{e_s}}{\Ets{\varphi(\vtheta_t^{s+1})}}
	+\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{\varphi(\vtheta_t^{s+1})}  \label{eq:new1}\\
	&= 
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}} \nonumber\\
	&\qquad+
	\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{\varphi(\vtheta_t^{s+1})} \label{eq:8}
\end{align}
Hence:
\begin{align}
	\left|\Ets{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}}
	\right.&-\left.\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}}\right| 
	=
	\left|\Covts[t-1]{\gradApp{\tilde{\vtheta}^s}{N}}{\varphi(\vtheta_t^{s+1})}\right|  
	\nonumber\\
	&\leq
	\sqrt{\Vars{\gradApp{\tilde{\vtheta}^s}{N}}}\circ\sqrt{\Varts[t-1]{\varphi(\vtheta_t^{s+1})}} \nonumber\\
	%
	&\leq	
	\frac{1}{2}\Vars{\gradApp{\tilde{\vtheta}^s}{N}} +\frac{1}{2}\Varts[t-1]{\varphi(\vtheta_t^{s+1})}\label{eq:9}\\
	%
	&=
	\frac{1}{2N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} +\frac{1}{2}\Varts[t-1]{\varphi(\vtheta_t^{s+1})} \label{eq:10}\\
	%
	&\leq
	\frac{1}{2N}\Vars{g(\cdot\vert\tilde{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{\varphi(\vtheta_t^{s+1})}^2},
	\nonumber
\end{align}
where (\ref{eq:6}) is from Definition \ref{def:ideal}; (\ref{eq:7}) is from the fact that $\gradIdeal{\vtheta_t^{s+1}}$ is both unbiased and independent from $\varphi(\vtheta_t^{s+1})$ \wrt the sampling at time $t$ alone, which is not true for $\gradBlack{\vtheta_t^{s+1}}$; 
(\ref{eq:new1}) is from the fact that $\gradJ{\tilde{\vtheta}^s}$ is constant \wrt $\Vars{\cdot}$;
(\ref{eq:8}) is from $\Ets{e_s}=0$; (\ref{eq:9}) is from Young's inequality; and (\ref{eq:10}) is from the definition of $\gradApp{\vtheta}{N}$.
\end{proof}

>>>>

We can now introduce Stochastic Variance-Reduced Policy Gradient (SVRPG) for a generic policy gradient estimator $g$ (see Eq.~\eqref{E:policygradient.estimate}).
% For simplicity, we focus on the REINFORCE estimator but the extension to G(PO)MDP is straightforward.
Between snapshots, the SVRPG estimate is corrected using importance weighting, for any $t \in \{0,\ldots,m-1\}$
\begin{align*}
        \blacktriangledown J(\vtheta_{t}) &= \wh{\nabla}_N J(\wt{\vtheta}) + g(\tau_i|\vtheta_t) - \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta}),
\end{align*}
% \begin{align*}
% 	\blacktriangledown J(\vtheta) &= \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \vert \wt{\vtheta})\Reward(\tau_j) \\
% 		&+ \nabla\log\pi(\tau_i \vert \vtheta)\Reward(\tau_i) 
% 		- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \wt{\vtheta})\Reward(\tau_i),
% \end{align*}
where $\wh{\nabla}_N J(\wt{\vtheta})$ is as in Eq.~\eqref{E:policygradient.estimate} where $\mathcal{D}_N$ is sampled using the snapshot policy $\pi_{\wt{\vtheta}}$, $\tau_i$ is sampled from the current policy $\pi_{\vtheta_t}$, and $\omega(\tau|\vtheta_t, \wt{\vtheta}) = \frac{p(\tau|\wt{\vtheta})}{p(\tau|\vtheta_t)}$ is an importance weight from $\pi_{\vtheta_t}$ to the snapshot policy $\pi_{\wt{\vtheta}}$. 
Similarly to SVRG, we have that $\vtheta_0 := \wt{\vtheta}$, and the update is a FG step.
Our update is still fundamentally on-policy since the weighting concerns only the correction term. However, this partial ``off-policyness'' is enough to introduce variance. This is a well-known issue of importance sampling~\citep[\eg][]{thomas2015high}. To mitigate it, we use mini-batches of trajectories of size $B \ll N$ to average the correction, \ie
\begin{align}\label{E:svrpg.estimate.batch}
        \blacktriangledown J(\vtheta_{t}) &:= v_t= \wh{\nabla}_N J(\wt{\vtheta})\\ \notag
                                            & \quad{} + 
        \underbracket{
        \frac{1}{B} \sum_{i=0}^{B-1} \left[
        g(\tau_i|\vtheta_t) - \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})
\right]}_{c_t}.
\end{align}
We want just to stress that $\EVV[\tau_i \sim p(\cdot|\vtheta_t)]{\frac{1}{B} \sum_{i=0}^{B-1} \omega(\tau_i|\vtheta_t, \wt{\vtheta}) g(\tau_i|\wt{\vtheta})} = \nabla J(\wt{\vtheta})$.\footnote{The reader can refer to Appendix~\ref{} for off-policy gradients and variants of REINFORCE and G(PO)MDP.}
This property will be used to prove Lemma~\ref{L:svrpg.properties}.
%\begin{align*}
%&\blacktriangledown J(\vtheta) = \frac{1}{N}\sum_{j=0}^{N-1}\nabla\log\pi(\tau_j \vert \wt{\vtheta})\Reward(\tau_j) +\\
%&\quad\frac{1}{B}\left[\sum_{i=0}^{B-1}
%\nabla\log\pi(\tau_i \vert \vtheta)\Reward(\tau_i) 
%- \omega(\tau_i)\nabla\log\pi(\tau_i \vert \wt{\vtheta})\Reward(\tau_i)\right].
%\end{align*}
The use of mini-batches is a common practice in SVRG since it can yield a performance improvement even in the supervised case~\citep{harikandeh2015stopwasting,konevcny2016mini}. It is easy to show that the SVRPG estimator has the following, desirable properties (see Appendix~\ref{app:proofs}):
\begin{lemma}\label{L:svrpg.properties} \todopirout{extended to any unbiased estimator}
        {\color{blue} Let $\wh{\nabla}_N J(\vtheta)$ be an unbiased estimator of~\eqref{E:policygradient} }
%Denote by $\wh{\nabla}_N^{\textsc{RF}} J(\vtheta)$ the REINFORCE estimator in~\eqref{E:policygradient.estimate} 
and let $\vtheta^* \in \argmin_{\vtheta} \{J(\vtheta)\}$. Then, the SVRG estimate in~\eqref{E:svrpg.estimate.batch} is \emph{unbiased}
\[
\mathop{\mathbb{E}}
\left[\blacktriangledown J(\vtheta)\right] = \gradJ{\vtheta}.
\]
and regardless of the mini-batch size $B$:\footnote{
For any vector $\mathbf{x}$, we use $\Var[\mathbf{x}]$ to denote the trace of the covariance matrix, \ie $\Tr\EVV[]{(\mathbf{x}-\EVV[]{\mathbf{x}})(\mathbf{x}-\EVV[]{\mathbf{x}})^T}$.}
\[
	\Var\left[\gradBlack{\vtheta^*}\right] = 
    \Var\left[\wh{\nabla}_N J(\vtheta^*)\right].
\]
\end{lemma}

@@

\begin{algorithm}[tb]
        \caption{SVRPG}
	\label{alg:svrpg}
	\begin{algorithmic}
        \STATE {\bfseries Input:} number of epochs $S$, epoch size $m$, step size $\alpha$, batch size $N$, mini-batch size $B$, gradient estimator $g$, initial parameter $\vtheta_{0}^m := \wt{\vtheta}^0$
		\FOR{$s=0$ {\bfseries to} $S-1$}
        \STATE $\vtheta_0^{s+1} := \wt{\vtheta}^{s} = \vtheta_{m}^s$
        \STATE Sample $N$ trajectories $\{\tau_j\}$ from $p(\cdot\vert\wt{\vtheta}^{s})$
        \STATE $ \wt{\mu} = \gradApp{\wt{\vtheta}^{s}}{N}$ (see Eq.~\eqref{E:policygradient.estimate})% = \frac{1}{N}\sum_{j=0}^{N-1}g(\tau_j | \wt{\vtheta}^s)$
		\FOR{$t=0$ {\bfseries to} $m-1$}
        \STATE Sample $B$ trajectories $\{\tau_i\}$ from $p(\cdot\vert\vtheta_t^{s+1})$
%        \STATE $c^{s+1}_t = \frac{1}{B} \sum\limits_{i=0}^{B-1} \left( g(\tau_i|\vtheta_t^{s+1}) - \omega(\tau_i|\vtheta^{s+1}_t, \wt{\vtheta}^s) g(\tau_i| \wt{\vtheta}^s) \right)$
        \STATE $c^{s+1}_t = \frac{1}{B} \sum\limits_{i=0}^{B-1}
        \begin{aligned}[t]
                \Big( & g(\tau_i|\vtheta_t^{s+1})\\ 
                      & \quad{} - \omega(\tau_i|\vtheta^{s+1}_t, \wt{\vtheta}^s) g(\tau_i| \wt{\vtheta}^s) \Big)
        \end{aligned}$
%        \STATE $\begin{aligned}[t]
%                &v^{s+1}_t = \wt{\mu}\\ 
%                & +\frac{1}{B} \sum\limits_{i=0}^{B-1} \left( g(\tau_i|\vtheta_t^{s+1}) - \omega(\tau_i|\vtheta^{s+1}_t, \wt{\vtheta}^s) g(\tau_i| \wt{\vtheta}^s) \right)
%                \end{aligned}$
        \STATE $v^{s+1}_t = \wt{\mu} + c^{s+1}_t$ % \gradApp{\wt{\vtheta}^{s}}{N} + c^{s}_t$
%%		\begin{align*}
%%		\blacktriangledown J(\vtheta_t^{s+1}) = 
%%		&\gradApp{\wt{\vtheta}^s}{N} \\
%%		&+\frac{1}{B}\sum_{i=0}^{B-1}\left[ 
%%		\score{}{\tau_i\vert\vtheta_t^{s+1}}\Reward(\tau_i)\right. \\
%%		&\left. - \omega(\tau_i)\score{}{\tau_i \vert \wt{\vtheta}^{s}}\Reward(\tau_i)\right]
%%		\end{align*}
        \STATE $\vtheta_{t+1}^{s+1} = \vtheta_t^{s+1} + \alpha v^{s+1}_t$
		\ENDFOR
		\ENDFOR
        \STATE {\bfseries return} $\vtheta_t^{s+1}$ with $(s,t)$ picked uniformly at random from $\{[0,S-1]\times[0,m-1]\}$
	\end{algorithmic}
\end{algorithm}

@@

\subsection{Baseline}
As reported in Section~\ref{subsec:PolicyGradient}, a popular variance reduction tool in Reinforcement Learning is the employment of baselines in policy gradient estimation.
This tool is orthogonal to the variance reduction technique proposed in this paper. In Section~\ref{sec:exp} we try to combine SVRPG with a non-stationary, state-dependent, learned linear baseline like the one used in~\cite{duan2016benchmarking}, with the variant that baseline parameters are updated only at the snapshot, with trajectories sampled from $\pi_{\wt{\vtheta}^s}$.

@@

+\frac{L_2^2}{B}\Ets[t-1]{\norm[2]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}
+\frac{1}{N}\Vars{g(\cdot\vert\wt{\vtheta}^s)}

@@

	= \Ets{\norm[]{\gradApp{\wt{\vtheta}^s}{N}

@@

    \omega(\tau_i)g(\tau_i\vert\wt{\vtheta}^s)}^2} \nonumber\\

@@

	&= \mathbb{E}_{t\vert s}\left[\left\|\gradApp{\wt{\vtheta}^s}{N}

@@

			\omega(\tau_i)g(\tau_i\vert\wt{\vtheta}^s)\right)

@@

            \pm \gradJ{\vtheta_t^{s+1}} \pm \gradJ{\wt{\vtheta}^s} 
			% -\gradJ{\vtheta_t^{s+1}} + \gradJ{\wt{\vtheta}^s}
			% +\gradJ{\vtheta_t^{s+1}} - \gradJ{\wt{\vtheta}^s}

@@

	+\Es{\norm[]{\gradApp{\wt{\vtheta}^s}{N} - \Es{\gradApp{\wt{\vtheta}^s}{N}}}^2} \nonumber\\

@@

				\omega(\tau_i)g(\tau_i\vert\wt{\vtheta}^s)\right)}\right\|^2\right] 

@@

	+\frac{1}{N}\Vars{g(\cdot\vert\wt{\vtheta}^s)}

@@

			\omega(\tau_i)g(\tau_i\vert\wt{\vtheta}^s)\right)}^2} \label{eq:2}\\

@@

    \omega(\tau_i)g(\tau_i\vert\wt{\vtheta}^s) \pm g(\tau_i|\wt{\vtheta}^s) }^2} \nonumber\\

@@

%%	+\frac{1}{N}\Vars{g(\cdot\vert\wt{\vtheta}^s)}

@@

%%			-g(\tau_i\vert\wt{\vtheta}^s)
%%			+g(\tau_i\vert\wt{\vtheta}^s) 
%%			-\omega(\tau_i)g(\tau_i\vert\wt{\vtheta}^s)}^2} \nonumber\\

@@

			-g(\tau_i\vert\wt{\vtheta}^s)}^2} 

@@

			\Ets{\norm[]{g(\tau_i\vert\wt{\vtheta}^s) 
			-\omega(\tau_i)g(\tau_i\vert\wt{\vtheta}^s)}^2} \nonumber\\

@@

    +\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}

@@

			-\omega(\tau_i))g(\tau_i\vert\wt{\vtheta}^s)}^2} \label{eq:3}\\

@@

	&\qquad+\frac{L_2^2}{B}\Ets[t-1]{\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}

@@

%Lemma 0
\begin{restatable}[]{lemma}{auxzero}\label{lemma:aux0}
Under Assumption \ref{ass:bounded_score}, for any function $\varphi(\vtheta_t^{s+1})$ which is deterministic for a fixed $\vtheta_t^{s+1}$:
\begin{align*}
\left|\Ets[t]{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}}
-\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}}
\right|
\leq
\frac{1}{2N}\Vars{g(\cdot\vert\wt{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{\varphi(\vtheta_t^{s+1})}^2}
\end{align*}
\todopir{I have to check this one!!!}
\end{restatable}
\begin{proof}
\begin{align}
	\Ets{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}}
	&=
	\Ets{\dotprod{\gradIdeal{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}} +
	\Ets[t-1]{\dotprod{e_s}{\varphi(\vtheta_t^{s+1})}} \label{eq:6}\\
	&=
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}} +
	\Ets[t-1]{\dotprod{e_s}{\varphi(\vtheta_t^{s+1})}} \label{eq:7}\\
	&=
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}} \nonumber\\
	&\qquad+
	\dotprod{\Ets{e_s}}{\Ets{\varphi(\vtheta_t^{s+1})}}
	+\Covts[t-1]{\gradApp{\wt{\vtheta}^s}{N}}{\varphi(\vtheta_t^{s+1})}  \label{eq:new1}\\
	&= 
	\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}} \nonumber\\
	&\qquad+
	\Covts[t-1]{\gradApp{\wt{\vtheta}^s}{N}}{\varphi(\vtheta_t^{s+1})} \label{eq:8}
\end{align}
Hence:
\begin{align}
	\left|\Ets{\dotprod{\gradBlack{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}}
	\right.&-\left.\Ets{\dotprod{\gradJ{\vtheta_t^{s+1}}}{\varphi(\vtheta_t^{s+1})}}\right| 
	=
	\left|\Covts[t-1]{\gradApp{\wt{\vtheta}^s}{N}}{\varphi(\vtheta_t^{s+1})}\right|  
	\nonumber\\
	&\leq
	\sqrt{\Vars{\gradApp{\wt{\vtheta}^s}{N}}}\circ\sqrt{\Varts[t-1]{\varphi(\vtheta_t^{s+1})}} \nonumber\\
	%
	&\leq	
	\frac{1}{2}\Vars{\gradApp{\wt{\vtheta}^s}{N}} +\frac{1}{2}\Varts[t-1]{\varphi(\vtheta_t^{s+1})}\label{eq:9}\\
	%
	&=
	\frac{1}{2N}\Vars{g(\cdot\vert\wt{\vtheta}^s)} +\frac{1}{2}\Varts[t-1]{\varphi(\vtheta_t^{s+1})} \label{eq:10}\\
	%
	&\leq
	\frac{1}{2N}\Vars{g(\cdot\vert\wt{\vtheta}^s)} +\frac{1}{2}\Ets[t-1]{\norm[]{\varphi(\vtheta_t^{s+1})}^2},
	\nonumber
\end{align}
where (\ref{eq:6}) is from Definition \ref{def:ideal}; (\ref{eq:7}) is from the fact that $\gradIdeal{\vtheta_t^{s+1}}$ is both unbiased and independent from $\varphi(\vtheta_t^{s+1})$ \wrt the sampling at time $t$ alone, which is not true for $\gradBlack{\vtheta_t^{s+1}}$; 
(\ref{eq:new1}) is from the fact that $\gradJ{\wt{\vtheta}^s}$ is constant \wrt $\Vars{\cdot}$;
(\ref{eq:8}) is from $\Ets{e_s}=0$; (\ref{eq:9}) is from Young's inequality; and (\ref{eq:10}) is from the definition of $\gradApp{\vtheta}{N}$.
\end{proof}

%Lemma 1
\begin{restatable}[]{lemma}{auxone}\label{lemma:aux1}
Under Assumptions \ref{ass:bounded_score} ans \ref{ass:M2}, the expected squared norm of the true gradient $\gradJ{\vtheta_t^{s+1}}$, for appropriate choices of $\alpha_t\geq0$ and $\beta_t>0$, can be bounded as follows:
\[
	\Ets{\norm[]{\gradJ{\vtheta_t^{s+1}}}^2} \leq
	\frac{R_{t+1}^{s+1} - R_t^{s+1}}{\Psi_t} + \frac{d_tV}{N\Psi_t}
	+\frac{f_tW}{B\Psi_t},
\]
	where
\begin{align*}
	&R_t^{s+1}\coloneqq \Ets{J(\vtheta_t^{s+1}) - c_t\norm[]{\vtheta_t^{s+1}-\wt{\vtheta}^s}^2}, \\
	&c_{m-1} = 0, \\
	&c_t = c_{t+1}\left(1+\alpha_t\beta_t+\alpha_t+\frac{\alpha_t^2L^2}{B}\right)+\frac{\alpha_t^2L^3}{2B}, \\
	&\Psi_t = \alpha_t\left(\frac{1}{2}-\frac{c_{t+1}}{\beta_t}-\frac{\alpha_tL}{2}-\alpha_tc_{t+1}\right), \\
	&d_t = \frac{\alpha_t}{2}\left(1+2c_{t+1}+\alpha_tL+2\alpha_tc_{t+1}\right), \\
	&f_t = \alpha_t^2\frac{\Gamma(L+2c_{t+1})}{2},
\end{align*}
where $L=\max\left\{L1,L2\right\}$, \ie the greater of the Lipschitz constants from Lemmas \ref{lemma:lsmooth} and \ref{lemma:gsmooth}.
